---
title: "Simulation_SaffranAslinNewport96"
output:
  pdf_document: default
  html_document:
    df_print: paged
  html_notebook:
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/rohrlich/go/src/github.com/ccnlab/lang-acq/data/Saffran/")
```
### Load Libraries
```{r}
library(Matrix)
library(lme4)
library(emmeans)
library(ggplot2)
library(dplyr)
library(psych)
```
### Terminology
In-Word - the transition between syllables within a word - 100% predictable for the data in this simulation

Next-Word - the transition between syllables within a word - 33% predictable for the data in this simulation

### Factors
run - equivalent to a subject as each run starts with a different random seed - treated as a random effect

phase - pretrain/test or train/test

epoch - each epoch we average the values from the trial data

layer - a representation of the neurons of a brain area 

condition - whether the syllable transition is "in-word" or "to next-word"

cosine - the cosine value for the run/epoch/condition (the independent variable)

## ---------------------------------------------
### Import
### Pretraining Test Data
```{r}
saffranPretrain.df <- read.table("roh002319/WordSeg_0-24_preTstEpcTidy.tsv", header=FALSE, sep="\t")
saffranPretrain.df <- setNames(saffranPretrain.df, c("run","phase","epoch","layer","condition", "cosine"))
```
### Training Test Data
```{r}
saffranPostPretrain.df <- read.table("roh002319/WordSeg_0-24_tstEpcTidy.tsv", header=FALSE, sep="\t")
saffranPostPretrain.df <- setNames(saffranPostPretrain.df, c("run","phase","epoch","layer","condition", "cosine"))
```
### Pretraining and Training Test Data
```{r}
saffranPrePostTrain.df <- read.table("roh002319/WordSeg_0-24_prePostTstEpcTidy.tsv", header=FALSE, sep="\t")
saffranPrePostTrain.df <- setNames(saffranPrePostTrain.df, c("run","phase","epoch","layer","condition", "cosine"))
```
### Add a cosine dissimilarity column (1 - cosine)
```{r}
saffranPostPretrain.df$cosinedissim = 1 - saffranPostPretrain.df$cosine 
# View(saffranPrePostTrain.df)
```
### Pretraining - Boxplot condition X epoch
```{r}
v1 = c(1.5, 3.5, 5.5, 7.5, 9.5)
    v2 = c("10",  "20", "30", "40", "50")
    boxplot(cosine ~ condition*epoch, data=saffranPretrain.df,
        xlab = "Epoch",
        ylab = "Cosine Similarity",
        ylim = range(0.0:1.0),
        col = c("#3399ff", "#ff3333"), # websafe
        xaxt="n"
        )
legend("bottomright", legend = c("In-Word", "Next-Word"),
       col = c("#3399ff", "#ff3333"), # websafe
       pch=15, bty="n", pt.cex=2, horiz=F, inset=c(0.0, 0.80)
       )
axis(side=1, at = v1, labels=v2)
```
## ---------------------------------------------
## Pretrained Network
### Training (after pre) Boxplot condition X epoch
```{r}
v1 = c(1.5, 3.5, 5.5, 7.5, 9.5, 11.5, 13.5, 15.5, 17.5, 19.5)
    v2 = c("1",  "2", "3", "4", "5", "6", "7", "8", "9", "10")
    boxplot(cosine ~ condition*epoch, data=saffranPostPretrain.df,
        xlab = "Epoch",
        ylab = "Cosine Similarity",
        ylim = range(0.0:1.0),
        col = c("#3399ff", "#ff3333"), # websafe
        xaxt="n"
        )
legend("bottomright", legend = c("In-Word", "Next-Word"),
       col = c("#3399ff", "#ff3333"), # websafe
       pch=15, bty="n", pt.cex=2, horiz=F, inset=c(0.0, 0.0)
       )
abline(v=10.5, col="black", lty=2)
arrows(x0=12.5, y0=.8, x1=10.7, y1=.8, length=0.1)
text(x=12.7, y=.8, labels="End of infant training", pos=4)
axis(side=1, at = v1, labels=v2)
```
## ---------------------------------------------
### plot of pretest and test together
### Boxplot condition X epoch
```{r}
v1 = c(1.5, 3.5, 5.5, 7.5, 9.5, 11.5, 13.5, 15.5, 17.5, 19.5, 21.5, 23.5, 25.5, 27.5, 29.5)
    v2 = c("10",  "20", "30", "40", "50", "1",  "2", "3", "4", "5", "6", "7", "8", "9", "10")
    boxplot(cosine ~ condition*epoch, data=saffranPrePostTrain.df,
        xlab = "Epoch",
        ylab = "Cosine Similarity",
        ylim = range(0.0:1.0),
        col = c("#3399ff", "#ff3333"), # websafe
        xaxt="n"
        )
legend("bottomright", legend = c("In-Word", "Next-Word"),
       col = c("#3399ff", "#ff3333"), # websafe
       pch=15, bty="n", pt.cex=2, horiz=F, inset=c(0.0, 0.0)
       )
abline(v=10.5, col="black", lty=1)
text(x=5, y=.9, labels="pretraining", pos=1)
abline(v=20.5, col="black", lty=2)
arrows(x0=22.5, y0=.8, x1=20.7, y1=.8, length=0.1)
text(x=22.7, y=.8, labels="     end of\ninfant training", pos=4)
axis(side=1, at = v1, labels=v2, cex.axis=1.0)
```
### Boxplot condition X layer
```{r}
v1 = c(1.5, 3.5, 5.5, 7.5, 9.5)
v2 = c("CB", "CPB", "RB", "RPB", "STS")
boxplot(cosine ~ condition*layer, data=saffranPostPretrain.df,
        xlab = "Layer",
        ylab = "Cosine Similarity",
        ylim = range(0.0:1.0),
        xaxt="n",
        col = c("#3399ff", "#ff3333") # websafe
        )
legend("bottomright", legend = c("In-Word", "Next-Word"),
       col = c("#3399ff", "#ff3333"), # websafe
       pch=15, bty="n", pt.cex=2, horiz=F, inset=c(0.0, 0.0)
       )
axis(side=1, at = v1, labels=v2)
```
## Pretrained Network - Mixed Model Analysis
### Full Model with within factors epoch, layer, condition
```{r}
saffranPretrained.model_simple = lmer(cosine ~ epoch + layer + condition + (1|run), data=saffranPostPretrain.df, REML=FALSE)
summary(saffranPretrained.model_simple)
```
### Reduced Model - no condition factor
```{r}
saffranPretrained.model_nocondition = lmer(cosine ~ epoch + layer + (1|run), data=saffranPostPretrain.df, REML=FALSE)
summary(saffranPretrained.model_nocondition)
```
### Likelihood Ratio Test - full vs no condition factor (i.e. in-word vs next-word)
```{r}
anova(saffranPretrained.model_simple, saffranPretrained.model_nocondition)
```
### Reduced Model - no epoch factor
```{r}
saffranPretrained.model_noepoch = lmer(cosine ~ condition + layer + (1|run), data=saffranPostPretrain.df, REML=FALSE)
summary(saffranPretrained.model_noepoch)
```
### Likelihood Ratio Test - full vs no epoch factor
```{r}
anova(saffranPretrained.model_simple, saffranPretrained.model_noepoch)
```
### Full Model with condition x epoch interaction
```{r}
saffranPretrained.model_epoch_interaction = lmer(cosine ~ layer + epoch * condition + (1|run), data=saffranPostPretrain.df, REML=FALSE)
summary(saffranPretrained.model_epoch_interaction)
```
### Reduced Model - no layer factor
```{r}
saffranPretrained.model_nolayer = lmer(cosine ~ condition + epoch + (1|run), data=saffranPostPretrain.df, REML=FALSE)
summary(saffranPretrained.model_nolayer)
```
### Likelihood Ratio Test - full vs no layer factor
```{r}
anova(saffranPretrained.model_simple, saffranPretrained.model_nolayer)
```
## Pretrained Network - Interactions
### Full Model with condition x layer interaction
```{r}
saffranPretrained.model_interaction = lmer(cosine ~ epoch + layer * condition + (1|run), data=saffranPostPretrain.df, REML=FALSE)
summary(saffranPretrained.model_interaction)
```
### Likelihood Ratio Test - full additive vs full interaction
```{r}
anova(saffranPretrained.model_simple, saffranPretrained.model_interaction)
```
### Visualizing the Interaction (ggplot)
```{r}
p <- emmip(saffranPretrained.model_interaction,  condition ~ layer)
p + theme(
  panel.background = element_rect(fill = "white"),
  text = element_text(size = 14),
  axis.line = element_line(color = "black"),
  legend.position = c(.20, .85),
  legend.key = element_rect(fill = "white")
) +
  ylim(0.35, 0.60) +
  xlab("Layer") +
  ylab("Cosine Similarity") +
  scale_color_manual(values = c("#3399ff", "#ff3333"))
p
```
### Exploring the 2 axes of Auditory Cortex
```{r}
emm = emmeans(saffranPretrained.model_interaction, specs = ~ layer | condition)
summary(emm)

CB = c(1, 0, 0, 0, 0)
CPB = c(0, 1, 0, 0, 0)
RB = c(0, 0, 1, 0, 0)
RPB = c(0, 0, 0, 1, 0)
STS = c(0, 0, 0, 0, 1)

contrast(emm, method = list("STS - CPB" = STS - CPB,
                            "STS - RPB" = STS - RPB,
                            "CPB - CB" = CPB - CB,
                            "RPB - RB" = RPB - RB,
                            "RPB - CPB" = RPB - CPB,
                            "RB - CB" = RB - CB))
```

## --------------------------------------------
## Pretraining Data  - Mixed Model Analysis
### Full Model with within factors epoch, layer, condition
```{r}
saffranPre.model_simple = lmer(cosine ~ epoch + layer + condition + (1|run), data=saffranPretrain.df, REML=FALSE)
summary(saffranPre.model_simple)
```
### Reduced Model - no condition factor
```{r}
saffranPre.model_nocondition = lmer(cosine ~ epoch + layer + (1|run), data=saffranPretrain.df, REML=FALSE)
summary(saffranPre.model_nocondition)
```
### Pretrain - Likelihood Ratio Test - full vs no condition factor (i.e. in-word vs next-word)
```{r}
anova(saffranPre.model_simple, saffranPre.model_nocondition)
```
### Reduced Model - no epoch factor
```{r}
saffranPre.model_noepoch = lmer(cosine ~ layer + condition + (1|run), data=saffranPretrain.df, REML=FALSE)
summary(saffranPre.model_noepoch)
```
### Pretrain - Likelihood Ratio Test - full vs no epoch factor
```{r}
anova(saffranPre.model_simple, saffranPre.model_noepoch)
```
## Original Saffran Experiment 2 Comparison with Simulation
### Subset of Simulation Data: Epoch 5 Only (comparable training point)
```{r}
epoch5.df <- select(filter(saffranPostPretrain.df, epoch == 104),c(1:7))
epoch5stats <- aggregate(epoch5.df$cosinedissim, by = list(cond = epoch5.df$condition), FUN = function(x) c(mean = mean(x), sd = sd(x), n = length(x)))
epoch5stats <- do.call(data.frame, epoch5stats)
epoch5stats$se <- epoch5stats$x.sd / sqrt(epoch5stats$x.n)
```
### Epoch 5 Data Only - full additive model
```{r}
saffranPretrainedEpoch5.model_simple = lmer(cosinedissim ~ layer + condition + (1|run), data=epoch5.df, REML=FALSE)
summary(saffranPretrainedEpoch5.model_simple)
```
### Epoch 5 Data Only - reduced model - no condition factor
```{r}
saffranPretrainedEpoch5.model_nocondition = lmer(cosinedissim ~ layer + (1|run), data=epoch5.df, REML=FALSE)
summary(saffranPretrainedEpoch5.model_nocondition)
```
### Likelihood Ratio Test - Epoch 5 data only - simple vs reduced no condition
```{r}
anova(saffranPretrainedEpoch5.model_simple, saffranPretrainedEpoch5.model_nocondition)
```
### Saffran Experiment 2 and Simulation Side by Side
```{r}
par(mfrow=c(1,2))

# Saffran Experiment 2
time <- c(6.77, 7.60)
se <- c(0.44, 0.42)
barplot(time,
        ylim=c(0,11),
        col = c("#3399ff", "#ff3333"), # websafe
        names.arg = c("Whole-Word", "Part-Word"),
        ylab = "Mean Listening Time (sec)",
        xlab = "Infants"
)

wwbarctr = 0.70
pwbarctr = 1.90
len = .025
segments(wwbarctr, time[1] - se[1], wwbarctr, time[1] + se[1], lwd=2)
segments(wwbarctr - len, time[1] - se[1], wwbarctr + len, time[1] - se[1], lwd=2)
segments(wwbarctr - len, time[1] + se[1], wwbarctr + len, time[1] + se[1], lwd=2)
segments(pwbarctr, time[2] - se[2], pwbarctr, time[2] + se[2], lwd=2)
segments(pwbarctr - len, time[2] - se[2], pwbarctr + len, time[2] - se[2], lwd=2)
segments(pwbarctr - len, time[2] + se[2], pwbarctr + len, time[2] + se[2], lwd=2)

# Simulation
cs <- epoch5stats$x.mean
cosineSim <- c(epoch5stats$x.mean[1], epoch5stats$x.mean[2])
se <- c(epoch5stats$se[1], epoch5stats$se[2])
barplot(cosineSim,
        ylim=c(0.0, 0.7),
        col = c("#3399ff", "#ff3333"), # websafe
        names.arg = c("In-Word\n(Whole-Word)", "Next-Word\n(Part-Word)"),
        ylab = "1 - Cosine Similarity",
        xlab = "Simulation"
)
wwSimbarctr = 0.70
pwSimbarctr = 1.90
len = .025
segments(wwSimbarctr, cosineSim[1] - se[1], wwSimbarctr, cosineSim[1] + se[1], lwd=2)
segments(wwSimbarctr - len, cosineSim[1] - se[1], wwSimbarctr + len, cosineSim[1] - se[1], lwd=2)
segments(wwSimbarctr - len, cosineSim[1] + se[1], wwSimbarctr + len, cosineSim[1] + se[1], lwd=2)
segments(pwSimbarctr, cosineSim[2] - se[2], pwSimbarctr, cosineSim[2] + se[2], lwd=2)
segments(pwSimbarctr - len, cosineSim[2] - se[2], pwSimbarctr + len, cosineSim[2] - se[2], lwd=2)
segments(pwSimbarctr - len, cosineSim[2] + se[2], pwSimbarctr + len, cosineSim[2] + se[2], lwd=2)
```
<!-- ## Alternate Analysis with Layers split into Axes -->
<!-- ### Pretraining Test Data -->
<!-- ```{r} -->
<!-- saffranPretrainAlt.df <- read.table("roh002300/WordSeg_0-24_preTstEpcTidy.tsv", header=FALSE, sep="\t") -->
<!-- saffranPretrainAlt.df <- setNames(saffranPretrainAlt.df, c("run","phase","epoch","axis1","axis2", "condition", "cosine")) -->
<!-- ``` -->
<!-- ### Pretraining Test Data -->
<!-- ```{r} -->
<!-- saffranPostPretrainAlt.df <- read.table("roh002300/WordSeg_0-24_tstEpcTidy.tsv", header=FALSE, sep="\t") -->
<!-- saffranPostPretrainAlt.df <- setNames(saffranPostPretrainAlt.df, c("run","phase","epoch","axis1","axis2","condition", "cosine")) -->
<!-- ``` -->
<!-- ### Pretraining Test Data -->
<!-- ```{r} -->
<!-- saffranPrePostTrainAlt.df <- read.table("roh002300/WordSeg_0-24_prePostTstEpcTidy.tsv", header=FALSE, sep="\t") -->
<!-- saffranPrePostTrainAlt.df <- setNames(saffranPrePostTrainAlt.df, c("run","phase","epoch","axis1","axis2","condition", "cosine")) -->
<!-- ``` -->
<!-- ### Full Model with within factors epoch, layer, condition -->
<!-- ```{r} -->
<!-- saffranPretrainedAlt.model_simple = lmer(cosine ~ epoch + axis1 + axis2 + condition + (1|run), data=saffranPrePostTrainAlt.df, REML=FALSE) -->
<!-- summary(saffranPretrainedAlt.model_simple) -->
<!-- ``` -->
<!-- ### Reduced Model - no condition factor -->
<!-- ```{r} -->
<!-- saffranPretrainedAlt.model_nocondition = lmer(cosine ~ epoch + axis1 + axis2 + (1|run), data=saffranPrePostTrainAlt.df, REML=FALSE) -->
<!-- summary(saffranPretrainedAlt.model_nocondition) -->
<!-- ``` -->
<!-- ### Likelihood Ratio Test - full vs no condition factor (i.e. in-word vs next-word) -->
<!-- ```{r} -->
<!-- anova(saffranPretrainedAlt.model_simple, saffranPretrainedAlt.model_nocondition) -->
<!-- ``` -->
